High geomagnetic field intensity recorded by anorthosite xenoliths requires a strongly powered late Mesoproterozoic geodynamo

### 2022-04-05 PNAS_manuscript/PNAS_Revision_Letter.tex
Updates from Overleaf

This is a good point. We did not make it clear in the text. 500\textdegree C is an estimated lab unblocking temperature for the anorthosites based on thermal demagnetization of \cite{Zhang2021b} and the zero-field heating steps of this study. Considering the cooling model of \cite{Zhang2021a} and the effect of lowering unblocking temperature due to slow cooling \cite{Pullaiah1975a}, an estimate for the cooling time for the anorthosite from the Curie temperature of magnetite to the natural unblocking temperature could be longer than 1 kyr. To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate, we use 10 kyr for calculating the cooling rate correction factor. This leads to a cooling rate of 2.5\times10$^{-10}$. This is the value we use for calculating the paleointensity estimate value of 36\%.

@@

Thanks for pointing this out. This MAD selection criteria needed to be updated. The data in fact have much smaller scatter, it is just the selection criteria we implemented were more generous than necessary. A MAD selection of $<$10 would result in the same selection as the highest specimen MAD is around 8. Our table Figure S3 also presents the small scatter of MAD.

@@

According to \cite{Tauxe2004a} DANG tests whether the component selected is actually trend- ing toward the origin as it measures the angle between the best-fit direction that is free floating and the direction between the centre of mass of the data and the origin of the vector component diagram. We did not use MAD $>$ DANG as an additional selection criteria and prefer to use MAD $<$ 5 and DANG $<$ 10.

@@

\textit{Figure S2. Can you indicate which of these specimens come from ``passing" samples and which are from ``failing" samples? Also, I'm pretty sure the ZFC>FC being an indication of MD material has been demonstrated to be unreliable. (Sorry I can't find the relevant reference.)}

Labels for whether the sister specimen to the rock chips used for MPMS measurements have been added. Figure caption has been updated with the following sentence: ``Pass" or ``Fail" represents whether their sister specimens passed our paleointensity selection or not.

>>>>

This is a good point. We did not make it clear in the text. 500\textdegree C is an estimated lab unblocking temperature for the anorthosites based on thermal demagnetization of \cite{Zhang2021b} and the zero-field heating steps of this study. Considering the cooling model of \cite{Zhang2021a} and the effect of lowering unblocking temperature due to slow cooling \cite{Pullaiah1975a}, an estimate for the cooling time for the anorthosite from the Curie temperature of magnetite to the natural unblocking temperature could be longer than 1 kyr. To fully account for the range of various lower temperature bounds used for optimize the paleointensity estimate, we use 10 kyr for calculating the cooling rate correction factor. This leads to a cooling rate of $2.5\times10^{-10}$. This is the value we use for calculating the paleointensity estimate value of 36\%.

@@

Thanks for pointing this out. This MAD selection criteria needed to be updated. The data in fact have much smaller scatter, it is just the selection criteria we implemented were more generous than necessary. A MAD selection of $<$10 would result in the same selection as the highest specimen MAD is around 8. Our table Figure S3 presents these MAD values.

@@

DANG is the Deviation ANGle which is the angle between the free-floating best-fit direction and the direction between data center of mass and the origin of the vector component diagram  \cite{Tauxe2004a}. As a result, DANG is considered to assess whether the component selected is actually trending toward the origin and we consider our usage to be appropriate.

@@

\textit{Figure S2. Can you indicate which of these specimens come from ``passing" samples and which are from ``failing" samples? Also, I'm pretty sure the ZFC$>$FC being an indication of MD material has been demonstrated to be unreliable. (Sorry I can't find the relevant reference.)}

Labels for whether the sister specimen to the rock chips used for MPMS measurements have been added. Figure caption has been updated with the following sentence: ``Pass" or ``Fail" represents whether their sister specimens passed our paleointensity selection criteria or not. We are not aware of the demonstration of ZFC$>$FC as being unreliable. The papers of Carter-Stiglitz et al. nicely demonstrate the effect for magnetite and titanomagnetite. Jackson and Moskowitz (2021) cite this effect as giving insight into MD behavior which indicates that these experts still consider it to be a reliable diagnostic.
